All Your Graphs Belong to Us

### 2012-07-20 doc/asplos13/implementation.tex
comments on implementation 'workers'

All global memory operations are implemented with active
messages. Requesters construct a descriptor holding the current
task/worker ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

\paragraph{Delegates} Delegate operations are used for low-locality accesses and synchronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation. We limit delegate
operations to operate on objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations for the same address from multiple
requesters are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

@@

\comment{If anything, this is just a footnote. -M

All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfering
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

@@

Grappa tasks are 32-byte entities: a 64-bit function pointer plus
three 64-bit arguments. We use three arguments because tasks are
usually generated as part of a parallel loop decomposition, and thus
each task needs three kinds of data:
\begin{description}
\item[Function pointer] This 64-bit value indicates what routine
  should run. We configure the system nodes to disable address randomization, so that we can depend on the fact that function pointers are valid across process images.
\item[Private argument] This 64-bit value is typically used for task-specific
  data (e.g., a loop index).
\item[Shared argument] This 64-bit value is typically used for data shared
  between all tasks that are part of a loop.
\item[Synchronization argument] This 64-bit value is typically used to determine
  when all tasks that are part of a loop have finished. It is usually
  a global pointer to a synchronization object allocated at the core
  that spawned the task.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitrary 64-bit values in the runtime, and can be used
for any purpose.

Tasks are not allocated any execution resources until the scheduler
decides to run them; when this occurs, tasks are matched with {\em
  workers}. Each worker is simply a collection of status bits and a
stack, allocated at a particular core.

@@

Since the compiler sees all calls to the context switch routine, we
can save even less state. Our context switch routine appears to the
compiler as inline assembly; we declare all the registers we need
to save as ``clobbered'' by the inline assembly routine, and the
compiler will issue its own save and restore code as needed. This allows the
compiler to avoid saving any registers that are not used, or are used
for temporary values that are not needed after the context switch.
\comment{I highly doubt this is important in real apps beyond our kernels.  -M

@@

\TODO{What is a worker?  We don't really describe this.  -M}

@@

In order to mitigate the low message injection rate limits of commodity
networks, Grappa's communication stack has two layers.

@@

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. Commodity networks including Infiniband
achieves their peak injection bandwidth \emph{only} when the packet
sizes are relatively large---on the order of multiple kilobytes. The
reason for this discrepancy is the combination of overheads associated
with handling each packet (in terms of bytes that form the actual
Infiniband packet, processing time at the card and processing on the
CPU within the driver stack). Our measurements confirm manufactures
published data~\cite{infinibandbandwidth}, that with this packet size
the bisection bandwidth is only a small fraction, less than
~\checkme{5\%} of the peak bisection bandwidth.

In our experiments the vast majority of requests were smaller than 255
bytes \TODO{more precise? median?}, far too small to make efficient
use of the network. To make the best use of the network, we must
convert our small messages into large ones.

>>>>

All global memory operations are implemented with active
messages. Requesters construct a descriptor holding the current
worker \TODO{reorder this paragraph after defined} ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

\paragraph{Delegates} Delegate operations are used for low-locality accesses and synchronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation \TODO{this is not an
    inherent limitation, it would be better to discuss in terms of
    fence -BM}. We limit delegate
operations to operate on objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations for the same address from multiple
requesters are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

@@

\comment{If anything, this is just a footnote. -M
All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfering
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.}

@@

Grappa tasks are 32-byte entities: a 64-bit function pointer plus
three 64-bit arguments. We use three arguments because tasks are
usually generated as part of a parallel loop decomposition, and thus
each task needs three kinds of data.
\begin{description}
\item[Function pointer] This 64-bit value indicates what routine
  should run. We configure the system nodes to disable address randomization, so that we can depend on the fact that function pointers are valid across process images.
\item[Private argument] This 64-bit value is typically used for task-specific
  data (e.g., a loop index).
\item[Shared argument] This 64-bit value is typically used for data shared
  between all tasks that are part of a loop or to specify the number
  of loop iterations.
\item[Synchronization argument] This 64-bit value is typically used to determine
  when all tasks that are part of a loop have finished. It is usually
  a global pointer to a synchronization object allocated at the core
  that spawned the task.
\end{description}
While these are the most common uses of the three task arguments, they
are treated as arbitrary 64-bit values in the runtime, and can be used
for any purpose.

Tasks are not allocated any execution resources until the scheduler
decides to run them; when this occurs, tasks are matched with {\em
  worker} threads. Each worker is simply a collection of status bits and a
stack, allocated at a particular core.

@@

Since the compiler sees all calls to the context switch routine, we
can save even less state. Our context switch routine appears to the
compiler as inline assembly; we declare all the registers we need
to save as ``clobbered'' by the inline assembly routine, and the
compiler will issue its own save and restore code as needed. This allows the
compiler to avoid saving any registers that are not used, or are used
for temporary values that are not needed after the context switch.
\comment{I highly doubt this is important in real apps beyond our kernels.  -M
Why not? liveness is a common compiler analysis -BM

@@

@@

In order to mitigate the low message injection rate limits of commodity
networks, Grappa's communication stack has two layers: one for
user-level messages and one for network-level messages.

@@

Grappa's lower networking layer aggregates the upper layer's messages
to improve performance. Commodity networks including Infiniband
achieves their peak injection bandwidth \emph{only} when the packet
sizes are relatively large---on the order of multiple kilobytes. The
reason for this discrepancy is the combination of overheads associated
with handling each packet (in terms of bytes that form the actual
Infiniband packet, processing time at the card and processing on the
CPU within the driver stack). Our measurements confirm manufacturers
published data~\cite{infinibandbandwidth}, that with this packet size
the bisection bandwidth is only a small fraction, less than
~\checkme{5\%} of the peak bisection bandwidth.

In our experiments the vast majority of requests were smaller than
\checkme{255}
bytes \TODO{more precise? median?}, far too small to make efficient
use of the network. To make the best use of the network, we must
convert our small messages into large ones.
\TODO{reference to aggregation evaluation later in the paper}
